---
title: "Statewide rents-to-income ratio map"
author: "Arthur Gailes"
date: "`r Sys.Date()`"
format: 
  html:
    code-fold: true
editor_options: 
  chunk_output_type: console
---

Project Description: A statewide map comparing the median home price in every jurisdiction to metropolitan household income. Generally, three to five is considered to be a healthy rate---this interactive map will reveal potentially exclusionary jurisdictions.

*Note*: the title says rents to income, but the description specifies home price, so I did both. Rents are compared to monthly income.

```{r setup}
#| include: false
#| message: false
pacman::p_load(sf, dplyr, mapview, here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}
#| echo: false
place_data <- st_read("data/tidy/place_acs_data.gpkg", quiet = TRUE)
ca_geo <- st_read("data/tidy/ca_shp.gpkg", quiet = TRUE)
ca_metro_data <- st_read("data/tidy/metro_acs_data.gpkg", quiet = TRUE)
```

```{r}
#| echo: false
#| label: data-proc
place_data_labs <- place_data |> mutate(
  home_inc_ratio = round(home_inc_ratio, 1),
  home_inc_fac = cut(
    home_inc_ratio, 
    c(0, 1, 3, 5, Inf),
    paste(c("0-1", "1-3", "3-5", "5 or more"), "to 1")),
  rent_inc_pct_str = scales::percent(rent_inc_pct, accuracy = 0.1),
  rent_inc_lab = cut(
    rent_inc_pct,
    c(0, 0.1, 0.3, 0.4, Inf),
    c("0-9%", "10-29%", "30-39% (Burdened)", "40% or more (Severely Burdened)")
  )
)
```

### Home Values to Income

```{r}
mapview(
  place_data_labs, 
  zcol = 'home_inc_fac', 
  label = "home_inc_ratio",
  layer.name = "Ratio of Home Values to Income")
```

### Rents as a percentage of monthly income

```{r}
mapview(
  place_data_labs, 
  zcol = 'rent_inc_lab',
  label = "rent_inc_pct_str",
  layer.name = "Gross Rent % of Monthly Income")
```
