---
title: "Statewide rents-to-income ratio map"
author: "Arthur Gailes"
date: "5/22/2021"
format: 
  html:
    code-fold: true
---

Project Description: A statewide map comparing the median home price in every jurisdiction to metropolitan household income. Generally, three to five is considered to be a healthy rate---this interactive map will reveal potentially exclusionary jurisdictions.

*Note*: the title says rents to income, but the description specifies home price, so I did both. Rents are compared to monthly income.

```{r setup}
#| include: false
#| message: false
pacman::p_load(sf, dplyr, mapview, here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
```{r}
#| include: false

place_data <- st_read("data/tidy/place_acs_data.gpkg", quiet = TRUE)
ca_geo <- st_read("data/tidy/ca_shp.gpkg", quiet = TRUE)
ca_metro_data <- st_read("data/tidy/metro_acs_data.gpkg", quiet = TRUE)
```

### Home Values to Income
```{r}
mapview(place_data, zcol = 'home_inc_ratio')
```

### Rents as a percentage of monthly income
```{r}
mapview(place_data, zcol = 'rent_inc_pct')
```